---
title: "Gerson de Kleuver"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    storyboard: true
    theme: flatly
    highlight: monochrome
---

```{r preprocessing, message = FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(flexdashboard)
library(compmus)
library(spotifyr)
library(plotly)

LP <- get_playlist_audio_features("", "37i9dQZF1DWVi45nh2EuPP")
GP <- get_playlist_audio_features("", "4bx5c78CAquCWNE4tw1reY")
RP <- get_playlist_audio_features("", "5AHH67GYsljwoB1q6UGvWg")

```

Introduction
============

SideBar {.sidebar}
--------------

#### Data sets

[Magic Music EDM Gaming Playlist](https://open.spotify.com/playlist/5AHH67GYsljwoB1q6UGvWg?si=01c590f404924c96):

* Songs:

* Release date:

[Gaming playlist](https://open.spotify.com/playlist/4bx5c78CAquCWNE4tw1reY?si=8b1700cc937d4d1d):

* Songs:

* Release date:

[League of Legends Official Playlist](https://open.spotify.com/playlist/37i9dQZF1DWVi45nh2EuPP?si=2829a0729b984f8e)

* Songs:

* Release date:

Column
--------------
Gaming music is a broad term which describes a group of music that is often related to the playing of watching of video games. It as a category has grown inmensely, as music and playlists carrying such label is being uploaded every day in large quantities to youtube and spotify.

More text TBA




Data exploration {.storyboard}
===============
### Comparison between the playlists
```{r}
XP <-
  bind_rows(
    LP %>% mutate(playlist = "League of legends playlist"),
    GP %>% mutate(playlist = "Gaming playlist"),
    RP%>% mutate(playlist = "Magic Music EDM Gaming Playlist")
  )

```
``` {r}
AuPlot <- XP %>%
  ggplot(aes(x = playlist, y = acousticness, fill=playlist)) +
  geom_violin() +
  ggtitle("Acousticness of gaming playlists") + theme(legend.position = "none") 
ggplotly(AuPlot)
```
***

More text TBA


### Correlations between gaming playlists
````{r}

pointPLOT <- XP %>%
  ggplot(aes(x = valence, fill=playlist,color=playlist,y = danceability, alpha=energy)) +
  geom_point() +
  ggtitle("Correlations between gaming playlists") +
  facet_wrap(~playlist) + theme(legend.position = "none") +
  geom_smooth()
                                
ggplotly(pointPLOT)
```

***

More text TBA

### Chroma gram of Warriors
```{r}
Warriors <-
  get_tidy_audio_analysis("3f4fc8c8unrQeKecmUPEDR") %>%
  select(segments) %>%
  unnest(segments) %>%
  select(start, duration, pitches)

Warriors %>%
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) %>%
  compmus_gather_chroma() %>% 
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  theme_minimal() +
  scale_fill_viridis_c()
```
***
This is league of legends most iconic song.
TBA

### Self-Similarity Matrice of Taking over

```{r}
TO <-
  get_tidy_audio_analysis("7asFSf2pkWNEG3E5EuN1QR") %>%
  compmus_align(bars, segments) %>%
  select(bars) %>%
  unnest(bars) %>%
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "acentre", norm = "manhattan"
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "mean"
      )
  )
bind_rows(
  TO %>%
    compmus_self_similarity(pitches, "aitchison") %>%
    mutate(d = d / max(d), type = "Chroma"),
  TO %>%
    compmus_self_similarity(timbre, "euclidean") %>%
    mutate(d = d / max(d), type = "Timbre")
) %>%
  mutate() %>%
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

```




Conclusion
============

More text TBA