---
title: "Portifolio"
author: "Gerson de Kleuver"
date: "22/02/2021"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    storyboard: true
    theme: flatly
    highlight: monochrome
---

```{r preprocessing, message = FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(flexdashboard)
library(spotifyr)
library(plotly)

LP <- get_playlist_audio_features("", "37i9dQZF1DWVi45nh2EuPP")
GP <- get_playlist_audio_features("", "4bx5c78CAquCWNE4tw1reY")
RP <- get_playlist_audio_features("", "5AHH67GYsljwoB1q6UGvWg")


```

Introduction
============

SideBar {.sidebar}
--------------

More text TBA


Column
--------------
Gaming music is a broad term which describes a group of music that is often related to the playing of watching of video games. It as a category has grown inmensely, as music and playlists carrying such label is being uploaded every day in large quantities to youtube and spotify.

More text TBA




Data exploration {.storyboard}
===============
### Comparison between the playlists
```{r}
XP <-
  bind_rows(
    LP %>% mutate(playlist = "League of legends playlist"),
    GP %>% mutate(playlist = "Gaming playlist"),
    RP%>% mutate(playlist = "Magic Music EDM Gaming Playlist")
  )

```
``` {r}
AuPlot <- XP %>%
  ggplot(aes(x = playlist, y = acousticness, fill=playlist)) +
  geom_violin() +
  ggtitle("Acousticness of gaming playlists") + theme(legend.position = "none") 
ggplotly(AuPlot)
```
***

More text TBA


### Correlations between gaming playlists
````{r}

pointPLOT <- XP %>%
  ggplot(aes(x = valence, fill=playlist,color=playlist,y = danceability, alpha=energy)) +
  geom_point() +
  ggtitle("Correlations between gaming playlists") +
  facet_wrap(~playlist) + theme(legend.position = "none") +
  geom_smooth()
                                
ggplotly(pointPLOT)
```

***

More text TBA

### Chroma gram of league of legends

***
Chroma gram TBA

Conclusion
============

More text TBA